---
title: 'Risky Visual Search - Exploratory Analysis'
author: 'Dave Braun'
date: 'December 19 2022'
output:
  html_document:
    theme: flatly
    code_folding: hide
    df_print: paged
    toc: true
    includes:
      after_body: ../../../html/footer.html
      in_header: ../../../html/favicon.html
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(inputFile,
    encoding = encoding,
    output_file = 'index.html')})
---


This document was last updated at `r Sys.time()`.

```{r include = FALSE}
library(tidyverse)
library(ggsci)
library(gridExtra)
library(ggsci)
library(ggrepel)
```

## The story thus far

Ran an overall model looking at the impact of framing while controlling for many other variables. The reversed reflection effect remains significant. Looked only at the timing controlled condition to assess need for cognition. Found a really interesting crossover interaction such that the reversed reflection effect is only present for those high in need for cognition.


```{r}
d <- read.csv('../../data/main_timing.csv')
N <- length(unique(d$subject))
head(d)

e2_demo <- read.csv('../../../exp2/data/exp2_demo.csv')


```

Initial sample size is `r N`.
```{r}
d <- d %>% 
  mutate(direction = recode(direction, `Easier than Reference` = 'Easier',
                          `Harder than Reference` = 'Harder'))
```
  
### Preprocess NFC scale

```{r}
## preprocess NFC scale

reverse_items <- c(3,4,5,7,8,9,12,16,17)
response_code <- data.frame(response = c('neutral', 'somewhat_agree', 'somewhat_disagree', 'strongly_agree', 'strongly_disagree', ''), numeric = c(3, 4, 2, 5, 1, NA))
survey <- e2_demo %>% 
  gather(question, response, cog_effort_q1:cog_effort_q18) %>% 
  select(subject, question, response) %>% 
  inner_join(response_code) 

reverse_items <- c(3,4,5,7,8,9,12,16,17)
response_code <- data.frame(response = c('neutral', 'somewhat_agree', 'somewhat_disagree', 'strongly_agree', 'strongly_disagree', ''), numeric = c(3, 4, 2, 5, 1, NA))
survey <- e2_demo %>% 
  gather(question, response, cog_effort_q1:cog_effort_q18) %>% 
  select(subject, question, response) %>% 
  inner_join(response_code) 

parse_question <- function(q) {
  return(as.numeric(sub('.', '', unlist(strsplit(q, split='_'))[3])))
}

survey <- survey %>% 
  mutate(question = Vectorize(parse_question)(question)) %>% 
  mutate(numeric_final = ifelse(question %in% reverse_items, 6-numeric, numeric)) %>% 
  arrange(subject) %>% 
  select(-response, -numeric) %>% 
  rename(response = numeric_final) 

d <- survey %>% 
    filter(!is.na(response)) %>% 
    group_by(subject) %>% 
    summarize(cog_need = mean(response)) %>% 
    right_join(d) 

```


### Compute performance metric

```{r}
sp <- d %>% 
    group_by(subject) %>% 
    summarize(ssd = mean(ssd), search_rt = mean(search_rt), error = mean(error)) 
sp %>% 
    ggplot(aes(x = search_rt, y = error, label = subject)) + 
    geom_point(aes(color = ssd)) +
    geom_smooth(method='lm') + 
    scale_color_distiller() + 
    labs(
        x = 'Mean Visual Search Response Time (ms)',
        y = 'Mean Error Rate',
        color = 'Proportion Selection \nof Safe Deck'
    ) + 
    theme_bw() + 
    theme(axis.ticks = element_blank(),
          panel.grid = element_blank())
```
**Compute correlation**
```{r}
cor.test(sp$search_rt, sp$error)
```


```{r}
sp$search_rt_c <- scale(sp$search_rt)[,1]
sp$error_c <- scale(sp$error)[,1]

min(sp$search_rt_c)
min(sp$error_c)

sp <- sp %>% 
    mutate(performance = sqrt((-2 - error_c)^2 + (-2 - search_rt_c)^2))
sp$performance <- -sp$performance

sp %>% 
    ggplot(aes(x = reorder(subject, performance), y = performance)) + 
    geom_bar(stat = 'identity') + 
    theme(axis.text.x = element_blank())

best <- sp %>% 
    arrange(-performance) %>% 
    select(subject, performance) %>% 
    slice_head(n = 10) %>% 
    select(subject)
best <- best$subject

sp$best <- ifelse(sp$subject %in% best, 'Best', 'Other')

sp %>% 
    ggplot(aes(x = search_rt, y = error, label = subject)) + 
    geom_point(aes(color = best)) +
    geom_smooth(method='lm') + 
    scale_color_npg() + 
    labs(
        title = 'Validating the Performance Metric',
        x = 'Mean Visual Search Response Time (ms)',
        y = 'Mean Error Rate',
        color = 'Performance'
    ) + 
    theme_bw() + 
    theme(axis.ticks = element_blank(),
          panel.grid = element_blank())

sp$performance <- scale(sp$performance)

d <- sp %>% 
    select(subject, performance) %>% 
    inner_join(d)

```



## Include performance into choice model

```{r}
library(lme4)
library(car)

d$delay_e <- ifelse(d$delay == 'Enforced', -0.5, 0.5)
d$direction_e <- ifelse(d$direction == 'Easier', -0.5, 0.5)
d$n_stimuli_lag_c <- scale(d$n_stimuli_lag)[,1]
d$magnitude_e <- ifelse(d$magnitude == 'Moderate', -0.5, 0.5)
d$trial_c <- ifelse(d$delay == 'Enforced', 
                    scale(d[d$delay=='Enforced',]$trial)[,1],
                    scale(d[d$delay == 'Not Enforced',]$trial)[,1])
d <- d[!is.na(d$n_stimuli_lag),]
# 
# m1 <- glmer(ssd ~ direction_e * delay_e * performance + 
#                 (1 + direction_e * delay_e | subject), 
#             data = d, family = binomial(link = 'logit'),
#             control = glmerControl(optimizer = 'bobyqa'),
#             nAGQ = 1)
# 
# m2 <- glmer(ssd ~ direction_e * delay_e * performance + 
#                 (1 | subject) + 
#                 (0 + direction_e * delay_e | subject), 
#             data = d, family = binomial(link = 'logit'),
#             control = glmerControl(optimizer = 'bobyqa'),
#             nAGQ = 1)
# 
# m3 <- glmer(ssd ~ direction_e * delay_e * performance + 
#                 (1 | subject) + 
#                 (0 + direction_e + delay_e | subject) +  
#                 (0 + direction_e:delay_e | subject), 
#             data = d, family = binomial(link = 'logit'),
#             control = glmerControl(optimizer = 'bobyqa'),
#             nAGQ = 1)
# 
# m4 <- glmer(ssd ~ direction_e * delay_e * performance + 
#                 (1 | subject) + 
#                 (0 + direction_e + delay_e | subject),  
#             data = d, family = binomial(link = 'logit'),
#             control = glmerControl(optimizer = 'bobyqa'),
#             nAGQ = 1)
# 
# m5 <- glmer(ssd ~ direction_e * delay_e * performance + 
#                 (1 | subject) + 
#                 (0 + direction_e | subject) + 
#                 (0 + delay_e | subject),
#             data = d, family = binomial(link = 'logit'),
#             control = glmerControl(optimizer = 'bobyqa'),
#             nAGQ = 1)
# 
# m6 <- glmer(ssd ~ direction_e +
#                 trial_c + 
#                  magnitude_e +
#                 delay_e + 
#                 n_stimuli_lag_c + 
#                 performance + 
#                 (1 | subject) + 
#                 (0 + magnitude_e | subject) + 
#                 (0 + direction_e | subject) + 
#                 (0 + trial_c | subject) + 
#                 (0 + n_stimuli_lag_c | subject), 
#             data = d, family = binomial(link = 'logit'),
#             control = glmerControl(optimizer = 'bobyqa'),
#             nAGQ = 1)

m7 <- glmer(ssd ~ direction_e +
                trial_c + 
                 magnitude_e +
                delay_e + 
                n_stimuli_lag_c + 
                performance + 
                (1 | subject) + 
                (0 + direction_e | subject) + 
                (0 + trial_c | subject) + 
                (0 + magnitude_e | subject), 
            data = d, family = binomial(link = 'logit'),
            control = glmerControl(optimizer = 'bobyqa'),
            nAGQ = 1)
# anova(m6, m7)

# 
# m7 <- glmer(ssd ~ direction_e * delay_e * performance + 
#                 (1 | subject), 
#             data = d, family = binomial(link = 'logit'),
#             control = glmerControl(optimizer = 'bobyqa'),
#             nAGQ = 1)
# 
# anova(m7, m6)

mm <- m7
```   


```{r}


m_reduced <- glmer(ssd ~ direction_e + 
                       #magnitude_e + 
                       n_stimuli_lag_c + 
                       trial_c + 
                       delay_e + 
                       performance + 
                (1 | subject) + 
                (0 + magnitude_e | subject) + 
                (0 + trial_c | subject) + 
                (0 + direction_e | subject), 
            data = d, family = binomial(link = 'logit'),
            control = glmerControl(optimizer = 'bobyqa'),
            nAGQ = 1)

anova(mm, m_reduced)

summary(mm)
exp(fixef(mm))
exp(confint(mm, method = 'Wald'))
```
No interactions with n_stimuli on previous trial.

I think maybe just hit this in a quick sentence saying the effects hold when accounting for individual differences in task performance. Show in an appendix if needed.

```{r}
library(ggExtra)
ci <- exp(confint(mm, method = 'Wald'))
levels <- c('Intercept', 'Framing', 'Trial Count', 'Outcome Magnitude', 
            'Delay', 'Search Stimuli (t-1)', 'Performance')
               
parameter <- factor(levels, levels = rev(levels))
pd <- data.frame(parameter = parameter, estimate = exp(fixef(mm)),
                 lower = ci[5:(nrow(ci)),1], upper = ci[5:(nrow(ci)),2])

pd %>% 
    ggplot(aes(x = parameter, y = estimate)) + 
    geom_hline(yintercept = 1, linetype = 'dotted') + 
    geom_point(size = 2) + 
    geom_errorbar(aes(ymin = lower, ymax = upper), width = .5) + 
    coord_flip() + 
    labs(
        x = 'Parameter',
        y = 'Odds of Selecting Safe Deck'
    ) + 
    theme_bw() + 
    theme(axis.ticks = element_blank(),
          panel.grid = element_blank())



```
```{r}
r <- data.frame(mode = ranef(mm, condVar = TRUE))
r <- r[r$mode.term == 'direction_e',]
r$fixed <- fixef(mm)[2]
r$fixed_odds_ratio <- exp(r$fixed)
r$random_beta <- r$mode.condval + r$fixed
r$odds_ratio <- exp(r$random_beta)
r$sd_or <- exp(r$mode.condsd)

p1 <- r %>% 
    mutate(id = 1:(nrow(.))) %>% 
    ggplot(aes(x = reorder(id, odds_ratio), y = odds_ratio)) + 
    geom_hline(yintercept = 1, linetype = 'dotted') + 
    geom_hline(yintercept = unique(r$fixed_odds_ratio), linetype = 'dashed', color = 'blue') + 
    geom_point() + 
    ylim(0, 10) +
    labs(
        x = 'Participant',
        y = 'Increase in Odds of Selecting Safe Deck when Outcomes are Harder than Reference'
    ) + 
    coord_flip() +
    theme_bw() + 
    theme(
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank())

ggMarginal(p1, margins = 'x', fill = 'steelblue', alpha = .7)
```


Honestly inspecting the estimates this way is a bit sobering. Majority of subjects seem like they cluster more around the null hypothesis.


## Incorporating need for cognition

```{r}
e2 <- d[d$delay == 'Enforced',]
e2$cog_need_c <- scale(e2$cog_need)[,1]

e2 %>% 
    group_by(subject) %>% 
    summarize(cog_need = mean(cog_need_c), performance = -mean(performance)) %>% 
    ggplot(aes(x = performance, y = cog_need)) + 
    geom_point() + 
    geom_smooth(method = 'lm') + 
    labs(
        x = 'Standardized Task Performance Proficiency',
        y = 'Standardized Need for Cognition'
    ) + 
    theme_bw() + 
    theme(axis.ticks = element_blank(),
          panel.grid = element_blank())

```


```{r}
e2 <- d[d$delay == 'Enforced',]
e2$cog_need_c <- scale(e2$cog_need)[,1]

mm <- glmer(ssd ~ direction_e * cog_need_c +
                 performance +
                magnitude_e +
                trial_c + 
                n_stimuli_lag_c + 
                (1 | subject) + 
                (0 + magnitude_e | subject) + 
                (0 + trial_c | subject) + 
                (0 + direction_e | subject),
            data = e2,
            family = binomial(link = 'logit'),
            control = glmerControl(optimizer = 'bobyqa'),
            nAGQ = 1)

## test direction
print('Test main effect of direction')
m_direction <- glmer(ssd ~ direction_e:cog_need_c +
                cog_need_c + 
                 performance +
                magnitude_e +
                trial_c + 
                n_stimuli_lag_c + 
                (1 | subject) + 
                (0 + magnitude_e | subject) + 
                (0 + trial_c | subject) + 
                (0 + direction_e | subject),
            data = e2,
            family = binomial(link = 'logit'),
            control = glmerControl(optimizer = 'bobyqa'),
            nAGQ = 1)

anova(m_direction, mm)

## test direction X cog need
print('Test two way interaction of direction and need for cognition')
m_direction_cogneed <- glmer(ssd ~ direction_e +
                cog_need_c + 
                 performance +
                magnitude_e +
                trial_c + 
                n_stimuli_lag_c + 
                (1 | subject) + 
                (0 + magnitude_e | subject) + 
                (0 + trial_c | subject) + 
                (0 + direction_e | subject),
            data = e2,
            family = binomial(link = 'logit'),
            control = glmerControl(optimizer = 'bobyqa'),
            nAGQ = 1)

anova(mm, m_direction_cogneed)



summary(mm)
exp(fixef(mm))
exp(confint(mm, method = 'Wald'))

```

### Following up on direction x cog need

**Observed data**

```{r}
npg_blue <- pal_npg('nrc')(2)[2]
npg_red <- pal_npg('nrc')(2)[1]

e2 %>% 
    group_by(subject, direction) %>% 
    summarize(ssd = mean(ssd), cog_need = mean(cog_need_c)) %>% 
    ggplot(aes(x = cog_need, y = ssd)) + 
    geom_point(aes(color = direction)) + 
    geom_smooth(method = 'lm', aes(color = direction, fill = direction)) + 
    labs(
        x = 'Standardized Need for Cognition',
        y = 'Proportion Selection of Safe Deck',
        color = 'Direction',
        fill = 'Direction'
    ) + 
    scale_color_manual(values = c(`Easier` = npg_blue, `Harder` = npg_red)) + 
    scale_fill_manual(values = c(`Easier` = npg_blue, `Harder` = npg_red)) + 
    theme_bw() + 
    theme(axis.ticks = element_blank(),
          panel.grid = element_blank(),
          strip.background = element_rect(color = 'black', fill = 'white'))
```


Can frame this as like high NFC people have more of an absence of aversion for the reference level and are more willing to accept the risk, as opposed to being more motivated to obtain the easiest outcome.

**Inference**

```{r}
print('Effect of NFC for easier than reference')
e2$is_easier <- ifelse(e2$direction == 'Easier', 0, 1)

m1 <- glmer(ssd ~ is_easier * cog_need_c +
                 performance +
                magnitude_e +
                trial_c + 
                n_stimuli_lag_c + 
                (1 | subject) + 
                (0 + magnitude_e | subject) + 
                (0 + trial_c | subject) + 
                (0 + direction_e | subject),
            data = e2,
            family = binomial(link = 'logit'),
            control = glmerControl(optimizer = 'bobyqa'),
            nAGQ = 1)

m_contrast <- glmer(ssd ~ is_easier  +
                is_easier:cog_need_c + 
                 performance +
                magnitude_e +
                trial_c + 
                n_stimuli_lag_c + 
                (1 | subject) + 
                (0 + magnitude_e | subject) + 
                (0 + trial_c | subject) + 
                (0 + direction_e | subject),
            data = e2,
            family = binomial(link = 'logit'),
            control = glmerControl(optimizer = 'bobyqa'),
            nAGQ = 1)
anova(m1, m_contrast)

exp(fixef(m1))
exp(confint(m1, method = 'Wald'))
```


```{r}
print('Effect of NFC for harder than reference')
e2$is_harder <- ifelse(e2$direction == 'Harder', 0, 1)

m1 <- glmer(ssd ~ is_harder * cog_need_c +
                 performance +
                magnitude_e +
                trial_c + 
                n_stimuli_lag_c + 
                (1 | subject) + 
                (0 + magnitude_e | subject) + 
                (0 + trial_c | subject) + 
                (0 + direction_e | subject),
            data = e2,
            family = binomial(link = 'logit'),
            control = glmerControl(optimizer = 'bobyqa'),
            nAGQ = 1)

m_contrast <- glmer(ssd ~ is_easier  +
                is_harder:cog_need_c + 
                 performance +
                magnitude_e +
                trial_c + 
                n_stimuli_lag_c + 
                (1 | subject) + 
                (0 + magnitude_e | subject) + 
                (0 + trial_c | subject) + 
                (0 + direction_e | subject),
            data = e2,
            family = binomial(link = 'logit'),
            control = glmerControl(optimizer = 'bobyqa'),
            nAGQ = 1)
anova(m1, m_contrast)

exp(fixef(m1))
exp(confint(m1, method = 'Wald'))
```

```{r}
print('Effect of framing for low NFC')
e2$cog_need_c_l <- e2$cog_need_c + 1

m1 <- glmer(ssd ~ direction_e * cog_need_c_l +
                 performance +
                magnitude_e +
                trial_c + 
                n_stimuli_lag_c + 
                (1 | subject) + 
                (0 + magnitude_e | subject) + 
                (0 + trial_c | subject) + 
                (0 + direction_e | subject),
            data = e2,
            family = binomial(link = 'logit'),
            control = glmerControl(optimizer = 'bobyqa'),
            nAGQ = 1)

m_contrast <- glmer(ssd ~ cog_need_c_l  +
                direction_e:cog_need_c_l + 
                 performance +
                magnitude_e +
                trial_c + 
                n_stimuli_lag_c + 
                (1 | subject) + 
                (0 + magnitude_e | subject) + 
                (0 + trial_c | subject) + 
                (0 + direction_e | subject),
            data = e2,
            family = binomial(link = 'logit'),
            control = glmerControl(optimizer = 'bobyqa'),
            nAGQ = 1)
anova(m1, m_contrast)

exp(fixef(m1))
exp(confint(m1, method = 'Wald'))
```

```{r}
print('Effect of framing for high NFC')
e2$cog_need_c_h <- e2$cog_need_c - 1

m1 <- glmer(ssd ~ direction_e * cog_need_c_h +
                 performance +
                magnitude_e +
                trial_c + 
                n_stimuli_lag_c + 
                (1 | subject) + 
                (0 + magnitude_e | subject) + 
                (0 + trial_c | subject) + 
                (0 + direction_e | subject),
            data = e2,
            family = binomial(link = 'logit'),
            control = glmerControl(optimizer = 'bobyqa'),
            nAGQ = 1)

m_contrast <- glmer(ssd ~ cog_need_c_h  +
                direction_e:cog_need_c_h + 
                 performance +
                magnitude_e +
                trial_c + 
                n_stimuli_lag_c + 
                (1 | subject) + 
                (0 + magnitude_e | subject) + 
                (0 + trial_c | subject) + 
                (0 + direction_e | subject),
            data = e2,
            family = binomial(link = 'logit'),
            control = glmerControl(optimizer = 'bobyqa'),
            nAGQ = 1)
anova(m1, m_contrast)

exp(fixef(m1))
exp(confint(m1, method = 'Wald'))
```
Reversed reflection effect only significant at 1 SD above mean on need for cognition.

**Prediction**

```{r}
library(jtools)
easyPredCI <- function(model,newdata=NULL,alpha=0.05) {
    ## baseline prediction, on the linear predictor (logit) scale:
    pred0 <- predict(model,re.form=NA,newdata=newdata)
    ## fixed-effects model matrix for new data
    X <- model.matrix(formula(model,fixed.only=TRUE)[-2],newdata)
    beta <- fixef(model) ## fixed-effects coefficients
    V <- vcov(model)     ## variance-covariance matrix of beta
    pred.se <- sqrt(diag(X %*% V %*% t(X))) ## std errors of predictions
    ## inverse-link function
    linkinv <- family(model)$linkinv
    ## construct 95% Normal CIs on the link scale and
    ##  transform back to the response (probability) scale:
    crit <- -qnorm(alpha/2)
    linkinv(cbind(conf.low=pred0-crit*pred.se,
                  conf.high=pred0+crit*pred.se))
}


testdata <- expand.grid(cog_need_c = seq(-2, 2, by = .001), direction_e = c(-.5, .5),
                        performance = 0, trial_c = 0, n_stimuli_lag_c = 0, magnitude_e = 0)

testdata$proba <- predict(mm, newdata = testdata, re.form = NA, type = 'response')
testdata <- cbind(testdata, easyPredCI(mm, newdata = testdata))

preds <- testdata %>% 
   # group_by(cog_need_c, direction_e) %>% 
   # summarize(proba = mean(proba), conf.low = mean(conf.low), conf.high = mean(conf.high)) %>% 
    mutate(direction = ifelse(direction_e == -0.5, 'Easier', 'Harder'))

p3 <- e2 %>% 
    group_by(subject, direction) %>% 
    summarize(ssd = mean(ssd), cog_need_c = mean(cog_need_c)) %>% 
    ggplot(aes(x = cog_need_c, y = ssd)) + 
    geom_hline(yintercept = 0.5, linetype = 'dotted') + 
    geom_point(aes(color = direction)) + 
    geom_ribbon(data = preds, 
                aes(x = cog_need_c, ymin = conf.low, ymax = conf.high, fill = direction, y = proba),
                alpha = .6) + 
    geom_line(data = preds,
              aes(x = cog_need_c, y = proba, color = direction),
              size = 1.5) + 
    labs(
        x = 'Standardized Need for Cognition',
        y = 'Probability of Selecting the Safe Deck',
        color = 'Framing',
        fill = 'Framing'
    ) +
    scale_fill_manual(values = c(`Easier` = npg_blue, `Harder` = npg_red)) + 
    scale_color_manual(values = c(`Easier` = npg_blue, `Harder` = npg_red)) + 
    theme_bw() + 
    theme(axis.ticks = element_blank(),
          panel.grid = element_blank(),
          legend.position = c(.15, .85))

#ggMarginal(p3, fill = pal_npg('nrc')(3)[3], type = 'histogram', binwidth = .10,
          # alpha = 1)
p3
```












































