---
title: "Experiment 3 Power Analysis"
author: "Dave Braun"
date: "8/5/2019"
output: 
  html_document:
    df_print: paged
    includes:
      after_body: ../../analysis/footer.html
      in_header: ../../analysis/favicon.html
knit:
  (function(inputFile, encoding) {
      rmarkdown::render(inputFile,
                    encoding = encoding,
                    output_file = 'index.html')})
---
```{r include = FALSE}
library(tidyverse)
library(data.table)
library(lme4)
library(ez)
```

*This document was last updated at `r Sys.time()`.*

This document is dedicated to the power analysis for Experiment 3. I'm taking a different approach here than I did for Experiments 1 and 2. Since Experiment 2 is so similar to this one, and I fit a GLMM to the data, which captures the fixed and random effects estimates from that data, I'll use simulations from that model to estimate the necessary sample for 0.8 power in Experiment 3.  

**TL;DR**: I'll want a sample size of 130 for Experiment 3.


```{r}
d <- read.csv('../../analysis/exp2/data/dstCleanChoice.csv')
d$selSafeDeck <- ifelse(d$selectedRiskyDeck == 1, 0, 1)
head(d)
```


## Fit the model to Experiment 2 data
I'm using the optimal model determined from [Experiment 2 Random Effects Analysis](../../analysis/exp2/scripts/exploratory/randomEffects/), which estimates a random intercept and two random slopes grouped by subject.

```{r}
d$differenceE <- ifelse(d$difference == 'Moderate', -0.5, 0.5)
d$difficultyE <- ifelse(d$difficulty == 'Easier than Reference', -0.5, 0.5)
m1 <- glmer(selSafeDeck ~ difficultyE * differenceE + (1 | subject) + (0 + difficultyE | subject) + (0 + differenceE | subject), data = d , family = binomial, control = glmerControl(optimizer = 'bobyqa'))
```

## Use fitted model to run simulations


```{r include = FALSE}
library(simr)
```


```{r}
sim_ext_subject <- extend(m1, along='subject', n = 150)
sim_ext_subject
```


```{r warning = FALSE, eval = FALSE}
powerResult <- powerCurve(sim_ext_subject, test=fcompare(selSafeDeck ~ differenceE), along = 'subject', breaks = seq(70,150,by = 10), nsim = 100)
powerResultDf <- data.frame(summary(powerResult))
save(powerResultDf, file = 'powerResultDf.Rdata')
```

```{r}
load('powerResultDf.Rdata')
```


```{r}
ggplot(powerResultDf, aes(x = factor(nlevels), y = mean, group = 1)) + 
  geom_line(linetype = 'dotdash', color = 'dodgerblue4') +
  geom_errorbar(aes(ymin = lower, ymax = upper), color = 'dodgerblue4') +
  geom_point(shape = 19, color = 'dodgerblue4', fill = 'dodgerblue4') +
  geom_hline(yintercept = .8, linetype = 'dashed') + 
  labs(
    x = 'Sample Size',
    y = 'Power to Detect Difficulty Effect',
    caption = 'Each estimate computed via 100 MCMC simulations'
  ) + 
  ylim(0,1) +
  theme_bw() +
  annotate('text', x = factor(70), y = .85, label = '0.8 Power')
```

That was so much easier than what I did for the first two experiments, and probably much more accurate.  

**I'll want a sample size of 130 for Experiment 3.**




















