---
title: "Experiment 1 Power Analysis"
author: "Dave Braun"
date: "2/17/2022"
output: 
  html_document:
    df_print: paged
    includes:
      after_body: ../analysis/footer.html
      in_header: ../analysis/favicon.html
knit:
  (function(inputFile, encoding) {
      rmarkdown::render(inputFile,
                    encoding = encoding,
                    output_file = 'index.html')})
---
```{r include = FALSE}
library(tidyverse)
library(data.table)
library(lme4)
library(ez)
```

*This document was last updated at `r Sys.time()`.*

This document is dedicated to the power analysis for Experiment 1. I'm using a model fitted to the data from the decks exp I ran on the py pool in the fall, and using simulations from that model to estimate power in the present study.

**TL;DR**: I'll want a sample size of 130 for Experiment 3.


```{r}
d <- read.csv('dstCleanChoice1.csv')
d$ssd <- ifelse(d$selectedRiskyDeck == 1, 0, 1)
head(d)
```


## Fit the model to Experiment 3 data
I'm using the optimal model from both E2 and E3 of my dissertation, which estimates a random intercept and two random slopes grouped by subject.

```{r}
d$differenceE <- ifelse(d$difference == 'Moderate', -0.5, 0.5)
d$difficultyE <- ifelse(d$difficulty == 'Easier than Reference', -0.5, 0.5)
m1 <- glmer(ssd ~ difficultyE * differenceE + (1 | subject) + (0 + difficultyE | subject) + (0 + differenceE | subject), data = d , family = binomial, control = glmerControl(optimizer = 'bobyqa'))
```

## Use fitted model to run simulations


```{r include = FALSE}
library(simr)
```


```{r}
sim_ext_subject <- extend(m1, along='subject', n = 150)
sim_ext_subject
```


```{r warning = FALSE, eval = FALSE}
## logic of the below line: https://humburg.github.io/Power-Analysis/simr_power_analysis.html
powerResult <- powerCurve(sim_ext_subject, test=fcompare(ssd ~ differenceE), along = 'subject', breaks = seq(70,150,by = 10), nsim = 100)
powerResultDf <- data.frame(summary(powerResult))
save(powerResultDf, file = 'powerResultDf.Rdata')
```

```{r}
load('powerResultDf.Rdata')
```


```{r}
ggplot(powerResultDf, aes(x = factor(nlevels), y = mean, group = 1)) + 
  geom_line(linetype = 'dotdash', color = 'dodgerblue4') +
  geom_errorbar(aes(ymin = lower, ymax = upper), color = 'dodgerblue4') +
  geom_point(shape = 19, color = 'dodgerblue4', fill = 'dodgerblue4') +
  geom_hline(yintercept = .8, linetype = 'dashed') + 
  labs(
    x = 'Sample Size',
    y = 'Power to Detect Difficulty Effect',
    caption = 'Each estimate computed via 100 MCMC simulations'
  ) + 
  ylim(0,1) +
  theme_bw() +
  annotate('text', x = factor(70), y = .85, label = '0.8 Power')
```

Alright...






```{r warning = FALSE, eval = FALSE}
## logic of the below line: https://humburg.github.io/Power-Analysis/simr_power_analysis.html
powerResult <- powerCurve(sim_ext_subject, test=fcompare(ssd ~ differenceE), along = 'subject', breaks = seq(10, 70,by = 10), nsim = 100)
powerResultDf <- data.frame(summary(powerResult))
save(powerResultDf, file = 'powerResultDf10-70.Rdata')
```
```{r}
load('powerResultDf10-70.Rdata')
```
```





```{r}
ggplot(powerResultDf, aes(x = factor(nlevels), y = mean, group = 1)) + 
  geom_line(linetype = 'dotdash', color = 'dodgerblue4') +
  geom_errorbar(aes(ymin = lower, ymax = upper), color = 'dodgerblue4') +
  geom_point(shape = 19, color = 'dodgerblue4', fill = 'dodgerblue4') +
  geom_hline(yintercept = .8, linetype = 'dashed') + 
  labs(
    x = 'Sample Size',
    y = 'Power to Detect Difficulty Effect',
    caption = 'Each estimate computed via 100 MCMC simulations'
  ) + 
  ylim(0,1) +
  theme_bw() +
  annotate('text', x = factor(70), y = .85, label = '0.8 Power')
```























