---
title: 'Risky Visual Search - Preprocessing'
author: 'Dave Braun'
date: December 13, 2022
output:
  html_document:
    theme: flatly
    code_folding: hide
    df_print: paged
    toc: true
    includes:
      after_body: ../../../html/footer.html
      in_header:  ../../../html/favicon.html
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(inputFile,
    encoding = encoding,
    output_file = 'index.html')})
---


This document was last updated at `r Sys.time()`.

```{r include = FALSE}
#Sys.setenv(RSTUDIO_PANDOC="/usr/lib/rstudio/pandoc/")
library(tidyverse)
```


```{r}
e1 <- read.csv('../../data/exp1a_main_raw.csv')
e2 <- read.csv('../../data/exp2_main_raw.csv')
e1$delay <- 'Not Enforced'
e2$delay <- 'Enforced'
d <- rbind(e1, e2)

N <- length(unique(d$subject))
head(d)
```


Initial sample size is `r N`.


## Check runtimes

(mostly for adjusting after pilot)

```{r}
d %>% 
  group_by(subject) %>% 
  summarize(runtime = max(phase_runtime_mins)) %>% 
  ggplot(aes(x = reorder(subject, runtime), y = runtime)) + 
  geom_bar(stat = 'identity') +
  geom_hline(yintercept = 30, linetype = 'dashed') +
  labs(
    x = 'Subject',
    y = 'Main Phase Runtime (mins)'
  ) + 
  theme(axis.text.x = element_blank())
```

Going to take everyone under an hour and average as the completion time.

```{r}
runtimes <- d %>% 
  group_by(subject) %>% 
  summarize(runtime = max(phase_runtime_mins)) 
avg_runtime <- mean(runtimes[runtimes$runtime < 60,]$runtime)


desired_trials <- function(avg_runtime, n_trials = 308) {
  mins_per_trial <- avg_runtime / n_trials
  return(30 / mins_per_trial)
}

print(paste0('Target trials to be 30 min: ', round(desired_trials(avg_runtime))))
```




## Accuracy & Response Time

For visual search

```{r}
## rt distribution
d %>% 
  filter(subject %in% sample(unique(d$subject), 10)) %>% 
  ggplot(aes(x = search_rt)) + 
  geom_histogram() + 
  facet_wrap(~subject, scales = 'free') + 
  labs(
    caption = 'A randomly-sampled ten participants'
  )


## calc proportion dropped
proportion_dropped <- function(rows_before, rows_now) {
  return((rows_before - rows_now) / rows_before)
}

d %>% 
  group_by(subject) %>% 
  summarize(accuracy = 1 - mean(error), rt = mean(search_rt)) %>% 
  ggplot(aes(x = rt/1000, y = accuracy, group = subject)) + 
  geom_hline(yintercept = .75, linetype = 'dotted') + 
  geom_vline(xintercept = .03, linetype = 'dotted') + 
  geom_point() + 
  labs(
    x = ' Average Response Time (s)',
    y = 'Average Accuracy'
  ) + 
  theme_bw()
```

## Location bias

```{r}
d$selected_deck_location_num <- ifelse(d$selected_deck_location == 'right', 1, -1)
d %>% 
  group_by(subject) %>% 
  summarize(sdln = mean(selected_deck_location_num)) %>% 
  ggplot(aes(x = subject, y = abs(sdln))) + 
  geom_hline(yintercept = .9, linetype = 'dotted') + 
  geom_point() + 
  coord_flip() + 
  ylim(0,1) +
  labs(
    x = 'Subject',
    y = 'Location Bias'
  ) + 
  theme_bw() + 
  theme(axis.text.y = element_blank(),
        axis.ticks = element_blank())
```

## Compute variables

```{r}
## compute lagged difficulty
d <- d %>% 
    group_by(subject) %>% 
    mutate(n_stimuli_lag = lag(n_stimuli)) 

```


## Data Trimming

```{r}
## cut subjects under 75% accuracy
bad_subjects_accuracy <- d %>% 
  group_by(subject) %>% 
  summarize(error = mean(error), rt = mean(search_rt), location_bias = abs(mean(d$selected_deck_location_num))) %>% 
  filter(error > .25 | rt < .03) %>% 
  select(subject)
n_bad_subjects_accuracy <- nrow(bad_subjects_accuracy)

bad_subjects_location <- d %>% 
  group_by(subject) %>% 
  summarize(location_bias = abs(mean(selected_deck_location_num))) %>% 
  filter(location_bias > .9) %>% 
  select(subject)
n_bad_subjects_location <- nrow(bad_subjects_location)

d <- d[!(d$subject %in% bad_subjects_accuracy$subject),]
d <- d[!(d$subject %in% bad_subjects_location$subject),]
```

There were `r n_bad_subjects_accuracy` subjects dropped for having less than 75% accuracy and `r n_bad_subjects_location` subjects dropped for having greater than 0.9 location bias.

### Gross level trial trimming

```{r}
## gross rt threshold
rows_before <- nrow(d)
d <- d[d$search_rt < 20000,]
trim_summary <- data.frame(reason = 'Search RTs above 20s', proportion_trimmed = proportion_dropped(rows_before, nrow(d)))
rows_before <- nrow(d)
d <- d[d$choice_rt < 20000,]
trim_summary <- rbind(trim_summary, data.frame(reason = 'Choice RTs above 20s', proportion_trimmed = proportion_dropped(rows_before, nrow(d))))

```


```{r}
## remove choice rts beyond 
hist(d$choice_rt)
```

### Fixed trimming on choice rt

```{r}
## Bounds on choice rt
rows_before <- nrow(d)
d <- d[d$choice_rt > 300 & d$choice_rt < 10000,]
print(paste0('Proportion dropped: ', round(proportion_dropped(rows_before, nrow(d)), 3), ', ', rows_before - nrow(d), '/', nrow(d)))
trim_summary <- rbind(trim_summary, data.frame(reason = 'Trimming choice rts less than 300 ms and greater than 10 s', proportion_trimmed = proportion_dropped(rows_before, nrow(d))))
```

### Trimming choice trials by participant-wise rt sds

```{r}
rows_before <- nrow(d)
d <- d %>% 
  group_by(subject) %>% 
  summarize(choice_rt_m = mean(choice_rt), choice_rt_sd = sd(choice_rt)) %>% 
  inner_join(d) %>% 
  filter(choice_rt >= choice_rt_m - (2*choice_rt_sd) & choice_rt <= choice_rt_m + (2*choice_rt_sd))
print(paste0('Proportion dropped: ', round(proportion_dropped(rows_before, nrow(d)), 3), ', ', rows_before - nrow(d), '/', nrow(d)))
trim_summary <- rbind(trim_summary, data.frame(reason = 'Trimming choice rts on +/- 2 participant wise SDs', proportion_trimmed = proportion_dropped(rows_before, nrow(d))))
```

```{r}
## code variables
d$ssd <- 1 - d$selectedRiskyDeck
```

### Randomly select 100 Ps

```{r}
set.seed(3329478)
final_subjects <- sample(unique(d$subject), size = 100)
d_final <- d[d$subject %in% final_subjects,]
```

```{r}
trim_summary
```



```{r}
## drop unnecessary columns
keep <- c('subject', 'trial', 'magnitude', 'direction', 'n_stimuli', 'n_stimuli_lag',
          'present_or_absent', 'delay', 'choice_rt', 'search_rt', 'error', 'ssd')
d_final <- d_final[, keep]

write.csv(d_final, '../../data/main_timing.csv', row.names = FALSE)
```

