---
title: 'Risky Visual Search - Experiment 1 and Experiment 3 Choice Analysis'
author: 'Dave Braun'
date: '`r format(Sys.time(), "%b %d %Y, %H:%M:%S")`'
output:
  html_document:
    theme: flatly
    code_folding: hide
    df_print: paged
    toc: true
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(inputFile,
    encoding = encoding,
    output_file = 'index.html')})
---


This document was last updated at `r Sys.time()`.

```{r include = FALSE}
library(tidyverse)
library(ez)
library(BayesFactor)
```


```{r}
e1 <- read.csv('../../../../../exp1a/data/exp1a_main.csv')
e3 <- read.csv('../../../../../exp3/data/exp3_main.csv')
N_e1 <- length(unique(e1$subject))
N_e3 <- length(unique(e3$subject))
N = N_e1 + N_e3
e1$experiment <- 'Experiment 1'
e3$experiment <- 'Experiment 3'
keep <- c('subject', 'magnitude', 'direction', 'ssd', 'experiment')
d <- rbind(e1[,keep], e3[,keep])
head(d)
```

Sample size for Experiment 1 is `r N_e1`.  
Sample size for Experiment 2 is `r N_e3`.  



```{r}
d %>% 
  group_by(subject, direction, magnitude, experiment) %>% 
  summarize(ssd_ = mean(ssd)) %>% 
  group_by(direction, experiment) %>% 
  summarize(ssd = mean(ssd_), se = sd(ssd_) / sqrt(N)) %>% 
  print() %>% 
  ggplot(aes(x = direction, y = ssd)) + 
  geom_bar(stat = 'identity') + 
  geom_errorbar(aes(ymin = ssd - se, ymax = ssd + se), width = .5) + 
  facet_wrap(~experiment) + 
  ylim(0,1) + 
  labs(
    x = 'Framing',
    y = 'Proportion Selection of Safe Deck',
    caption = paste0('N = ', N)
  ) + 
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))
```
```{r}
ggsave('figures/pdf/e1_e3_means.pdf', height = 5.5, width = 7.5)
ggsave('figures/png/e1_e3_means.png', height = 5.5, width = 7.5, dpi = 300, units = 'in')
```

### Individual differences in direction effect

```{r}
m <- d %>% 
  group_by(subject, direction, experiment) %>% 
  summarize(ssd_ = mean(ssd)) %>% 
  group_by(direction, experiment) %>% 
  summarize(ssd = mean(ssd_))

d %>% 
  group_by(subject, direction, experiment) %>% 
  summarize(ssd = mean(ssd)) %>% 
  ggplot(aes(x = direction, y = ssd)) + 
  geom_violin(fill = 'steelblue', alpha = .2) + 
  geom_boxplot(fill = NA) + 
  geom_jitter(width = .05, alpha = .7) + 
  geom_line(aes(group = subject), linetype = 'dashed', alpha = .7) +
  facet_wrap(~experiment) + 
  ylim(0,1) +
  labs(
    x = 'Direction',
    y = 'Proportion Selection of Safe Option',
    caption = 'N = 100'
  ) + 
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))

```



```{r}
ggsave('figures/pdf/e1_e3_individuals.pdf', height = 5.5, width = 7.5)
ggsave('figures/png/e1_e3_individuals.png', height = 5.5, width = 7.5, dpi = 300, units = 'in')
```


## Regular ANOVA

```{r}
library(ez)

ad <- d %>% 
  group_by(subject, experiment, direction, magnitude) %>% 
  summarize(ssd_ = mean(ssd)) 

m1 <- ezANOVA(wid = subject, within = .(direction, magnitude), between = .(experiment),
        dv = ssd_, detailed = TRUE, data = ad)

m1 <- cbind(m1$ANOVA, data.frame(n2p = m1$ANOVA$SSn / (m1$ANOVA$SSn + m1$ANOVA$SSd)))
m1
```

```{r}
d %>% 
  group_by(subject, direction) %>% 
  summarize(ssd_ = mean(ssd)) %>% 
  group_by(direction) %>% 
  summarize(ssd = mean(ssd_), se = sd(ssd_) / sqrt(N))
```

## Bayesian ANOVA

```{r}
## a useful resource: https://richarddmorey.github.io/BayesFactor/#mixed
d$direction <- factor(d$direction)
d$experiment <- factor(d$experiment)
d$magnitude <- factor(d$magnitude)
d$subject <- factor(d$subject)
b <- anovaBF(ssd ~ direction * magnitude * experiment + subject, data = d, whichRandom = 'subject')
b
```
Alternative / Null

More complex models

```{r}
b[17] / b[12]
```
Simpler models

```{r}
b[14] / b[8]
```
Or as testing for evidence in favor of the null (for simpler models):

```{r}
1 / 1.670621
```







## Dealing with the three-way

### Means and plot

I guess I need to reckon with all these crazy interactions.

```{r}
d %>% 
  group_by(subject, experiment, direction, magnitude) %>% 
  summarize(ssd_ = mean(ssd)) %>% 
  group_by(experiment, direction, magnitude) %>% 
  summarize(ssd = mean(ssd_), se = sd(ssd_) / sqrt(100)) %>% 
  print() %>% 
  mutate(magnitude = factor(magnitude, levels = c('Moderate', 'Extreme'))) %>% 
  ggplot(aes(x = magnitude, y = ssd, group = direction)) + 
  geom_hline(yintercept = .5, linetype = 'dotted') + 
  geom_bar(stat = 'identity', aes(fill = direction), position = position_dodge(width = .9), color = 'black') + 
  geom_errorbar(aes(ymin = ssd - se, ymax = ssd + se), position = position_dodge(width = .9), width = .5) + 
  facet_wrap(~experiment) + 
  labs(
    x = 'Magnitude',
    y = 'Proportion Selection of Safe Option',
    fill = 'Direction'
  ) + 
  scale_fill_manual(values = c(`Easier than Reference` = 'white', `Harder than Reference` = 'dark grey')) + 
  ylim(0,1) + 
  theme_bw() + 
  theme(strip.background = element_rect(fill = 'white', color = 'black'),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        legend.position = c(.2, .8))
```


### Statistics

Simple effect of `direction X magnitude` for Experiment 1.

```{r}
## full model is m1

formatSimpleEffects <- function(model, omnibusParams) {
  SSd <- omnibusParams['SSd']
  DFd <- omnibusParams['DFd']
  MSe <- SSd / DFd
  SSn <- model$ANOVA$SSn[nrow(model$ANOVA)]
  DFn <- model$ANOVA$DFn[nrow(model$ANOVA)]
  MSn <- SSn / DFn
  f_value <- MSn / MSe
  p_value <- pf(f_value, DFn, DFd, lower.tail = FALSE)
  n2p <- SSn / (SSn + SSd)
  return(paste('F(',DFn, ', ', omnibusParams['DFd'], ') = ', round(f_value, 2), ', p = ', round(p_value, 7), ', n2p = ', round(n2p, 3), sep = ''))
  
}

m_e1 <- ezANOVA(wid = subject, within = .(direction, magnitude), dv = ssd, detailed = TRUE, data = d[d$experiment == 'Experiment 1',])

omnibusParams <- c('SSd' = m1$SSd[nrow(m1)], 'MSe' = m1$SSd[nrow(m1)] / m1$DFd[nrow(m1)], 'DFd' = m1$DFd[nrow(m1)])
formatSimpleEffects(m_e1, omnibusParams)



```

Simple effect of `direction X magnitude` for Experiment 3



```{r}
m_e3 <- ezANOVA(wid = subject, within = .(direction, magnitude), dv = ssd, detailed = TRUE, data = d[d$experiment == 'Experiment 3',])

formatSimpleEffects(m_e3, omnibusParams)
```

**Within Experiment 3**

Simple effect of direction for moderate.

```{r}
m_e3_moderate <- ezANOVA(wid = subject, within = .(direction), dv = ssd, detailed = TRUE, data = d[d$experiment == 'Experiment 3' & d$magnitude == 'Moderate',])

formatSimpleEffects(m_e3_moderate, omnibusParams)
```
Simple effect of direction for extreme.

```{r}
m_e3_extreme <- ezANOVA(wid = subject, within = .(direction), dv = ssd, detailed = TRUE, data = d[d$experiment == 'Experiment 3' & d$magnitude == 'Extreme',])

formatSimpleEffects(m_e3_extreme, omnibusParams)
```






































