---
title: 'Risky Visual Search - Experiment 1 and Experiment 2 Choice Analysis'
author: 'Dave Braun'
date: '`r format(Sys.time(), "%b %d %Y, %H:%M:%S")`'
output:
  html_document:
    theme: flatly
    code_folding: hide
    df_print: paged
    toc: true
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(inputFile,
    encoding = encoding,
    output_file = 'index.html')})
---


This document was last updated at `r Sys.time()`.

Current plan (2022-06-08): Run the whole thing `direction X magnitude X experiment` both straight up and Bayesian style. See what kinda evidence the Bayes factor gives in the full model. Go from there.

```{r include = FALSE}
library(tidyverse)
library(ez)
library(BayesFactor)
```


```{r}
e1 <- read.csv('../../../../../exp1a/data/exp1a_main.csv')
e2 <- read.csv('../../../../../exp2/data/exp2_main.csv')
N_e1 <- length(unique(e1$subject))
N_e2 <- length(unique(e2$subject))
N = N_e1 + N_e2
keep <- c('subject', 'magnitude', 'direction', 'ssd')
e1$experiment <- 'Experiment 1'
e2$experiment <- 'Experiment 2'
d <- rbind(e1, e2)
head(d)
```

Sample size for Experiment 1 is `r N_e1`.  
Sample size for Experiment 2 is `r N_e2`.  



```{r}
d %>% 
  group_by(subject, direction, magnitude, experiment) %>% 
  summarize(ssd_ = mean(ssd)) %>% 
  group_by(direction, experiment) %>% 
  summarize(ssd = mean(ssd_), se = sd(ssd_) / sqrt(N)) %>% 
  print() %>% 
  ggplot(aes(x = direction, y = ssd)) + 
  geom_bar(stat = 'identity') + 
  geom_errorbar(aes(ymin = ssd - se, ymax = ssd + se), width = .5) + 
  facet_wrap(~experiment) + 
  ylim(0,1) + 
  labs(
    x = 'Framing',
    y = 'Proportion Selection of Safe Deck',
    caption = paste0('N = ', N)
  ) + 
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))
```
```{r}
ggsave('figures/pdf/e1_e2_means.pdf', height = 5.5, width = 7.5)
ggsave('figures/png/e1_e2_means.png', height = 5.5, width = 7.5, dpi = 300, units = 'in')
```

### Individual differences in direction effect

```{r}
m <- d %>% 
  group_by(subject, direction, experiment) %>% 
  summarize(ssd_ = mean(ssd)) %>% 
  group_by(direction, experiment) %>% 
  summarize(ssd = mean(ssd_))

d %>% 
  group_by(subject, direction, experiment) %>% 
  summarize(ssd = mean(ssd)) %>% 
  ggplot(aes(x = direction, y = ssd)) + 
  geom_violin(fill = 'steelblue', alpha = .2) + 
  geom_boxplot(fill = NA) + 
  geom_jitter(width = .05, alpha = .7) + 
  geom_line(aes(group = subject), linetype = 'dashed', alpha = .7) +
  facet_wrap(~experiment) + 
  ylim(0,1) +
  labs(
    x = 'Direction',
    y = 'Proportion Selection of Safe Option',
    caption = 'N = 100'
  ) + 
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))

```



```{r}
ggsave('figures/pdf/e1_e2_individuals.pdf', height = 5.5, width = 7.5)
ggsave('figures/png/e1_e2_individuals.png', height = 5.5, width = 7.5, dpi = 300, units = 'in')
```

# Assessing Whether Things Differ Across Experiments

## Regular ANOVA

```{r}
library(ez)

ad <- d %>% 
  mutate(ssd = 1 - selectedRiskyDeck) %>% 
  group_by(subject, experiment, direction, magnitude) %>% 
  summarize(ssd_ = mean(ssd)) 

m1 <- ezANOVA(wid = subject, within = .(direction, magnitude), between = .(experiment),
        dv = ssd_, detailed = TRUE, data = ad)

m1 <- cbind(m1$ANOVA, data.frame(n2p = m1$ANOVA$SSn / (m1$ANOVA$SSn + m1$ANOVA$SSd)))
m1
```

```{r}
d %>% 
  group_by(subject, direction) %>% 
  summarize(ssd_ = mean(ssd)) %>% 
  group_by(direction) %>% 
  summarize(ssd = mean(ssd_), se = sd(ssd_) / sqrt(N))
```

## Bayesian ANOVA

```{r}
## a useful resource: https://richarddmorey.github.io/BayesFactor/#mixed
d$direction <- factor(d$direction)
d$experiment <- factor(d$experiment)
d$magnitude <- factor(d$magnitude)
d$subject <- factor(d$subject)
b <- anovaBF(ssd ~ direction * magnitude * experiment + subject, whichRandom = 'subject', data = d)
b
```

Do I do this:

```{r}
b[15] / b[9]
```
or this

```{r}
b[17] / b[12]
```


I'm just curious to see what the Bayesian analysis has to say about the impact of direction only in E2.

```{r}
e2$direction <- factor(e2$direction)
e2$magnitude <- factor(e2$magnitude)
e2$subject <- factor(e2$subject)

e2_model <- anovaBF(ssd ~ direction * magnitude + subject, data = e2, whichRandom = 'subject')
e2_model

```
To flip a ratio, divide 1 over the ratio.

$$
a = \frac{b}{c}\\
d = \frac{c}{d}\\
d = \frac{c}{b} = \frac{1}{\frac{b}{c}} = \frac{1}{a}
$$

So main effect of direction evidence in favor of null:

```{r}
1 / exp(e2_model@bayesFactor$bf[2])
```


And the regular anova

```{r}
e2_reg <- ezANOVA(wid = subject, within = .(direction, magnitude), dv = ssd, data = e2, detailed = TRUE)
e2_reg$ANOVA
cbind(e2_reg$ANOVA, data.frame(n2p = e2_reg$ANOVA$SSn / (e2_reg$ANOVA$SSn + e2_reg$ANOVA$SSd)))
```





## Marginal means of difference for Experiment 2

```{r}
e2 %>% 
  group_by(subject, direction, magnitude) %>% 
  summarize(ssd_ = mean(ssd)) %>% 
  group_by(direction) %>% 
  summarize(ssd = mean(ssd_), se = sd(ssd_) / sqrt(100))
```






































